Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology

### 2021-06-30 MS/ms.Rmd
Adding some more detailed survey methods for categorisation / decisions. Its not easy, but hopfully this is more clear

As discussed above, it is clear that different types of questions and experimental approaches will be more or less susceptible to the 'apples and oranges' problem. <!--- not sure whether it is a good idea to highlight it as a problem as said earlier - explaining hetero is what's interesting and even the main purpose of meta-analysis for comparative physiologists - they are not the problem but they are what we are going to investigate. I do not think 'susceptible' is a good word here given the average hetero = > 90% - you are nearly 100% guaranteed to have hetero and this is what's interesting ---> As such, before embarking on our path to 'meta-analytic enlightenment', we conducted a literature survey to understand how widespread the problem may be in comparative physiology and gauge the ways in which it is currently being addressed, if at all. To accomplish this goal, we searched for meta-analyses and comparative analyses that tackled questions using physiological data compiled from animal studies across multiple species. We were specifically interested in asking: 1) what type of effect size was used within the quantitative synthesis? 2) based on the methods provided, was the effect size likely susceptible to variation in how treatments are applied across studies? 3) If so, was the variation resulting from treatment differences accounted for in the effect size or during analysis?, and finally, 4) how were effect sizes analysed in the quantitative synthesis, were these weighted analyses, and what program was used for analysis? We also recorded whether publication bias and sensitivity analyses were conducted, if data were publicly available, where it was stored, and if so, if it was usable (i.e., in a machine readable format).

@@

All title, abstract and full text screening was done by two people (DWAN and ML). Any conflicts or disagreements were resolved by discussion and a consensus on a paper was reached. Out of the 417 papers, 79 full texts were screened and we included a total of 62 for final data extraction. For full details on the search strings used, dates searched, the number of papers found, and our PRISMA flowchart (along with reasons for exclusion at the full text stage) we refer readers to the *Supplementary Materials*.

>>>>

As discussed above, it is clear that different types of questions and experimental approaches will be more or less susceptible to the 'apples and oranges' problem. <!--- not sure whether it is a good idea to highlight it as a problem as said earlier - explaining hetero is what's interesting and even the main purpose of meta-analysis for comparative physiologists - they are not the problem but they are what we are going to investigate. I do not think 'susceptible' is a good word here given the average hetero = > 90% - you are nearly 100% guaranteed to have hetero and this is what's interesting ---> As such, before embarking on our path to 'meta-analytic enlightenment', we conducted a literature survey to understand how widespread the problem may be in comparative physiology and gauge the ways in which it is currently being addressed, if at all. To accomplish this goal, we searched for meta-analyses and comparative analyses that tackled questions using physiological data compiled from animal studies across multiple species. We were specifically interested in asking: 1) what type of effect size was used within the quantitative synthesis? 2) based on the methods provided, was the effect size likely susceptible to continuous variation in how treatments were applied (e.g., for contarst-based effects) or the conditions under which traits were measured (i.e., for correlatioons) across studies? 3) If so, was the variation resulting from treatment differences accounted for in the effect size or during analysis?, and finally, 4) how were effect sizes analysed in the quantitative synthesis, were these weighted analyses, and what program was used for analysis? We also recorded whether publication bias and sensitivity analyses were conducted, if data were publicly available, where it was stored, and if so, if it was usable (i.e., in a machine readable format).

@@

All title, abstract and full text screening was done by two people (DWAN and ML). Any conflicts or disagreements were resolved by discussion and a consensus on a paper was reached. Out of the 417 papers, 79 full texts were screened and we included a total of 62 for final data extraction. For full details on the search strings used, dates searched, the number of papers found, and our PRISMA flowchart (along with reasons for exclusion at the full text stage) we refer readers to the *Supplementary Materials*. To address our questions, we categorised effect sizes as falling into the following categories: Response Ratio (lnRR), Standardised Mean Differences (SMD; e.g., Hedges d, g or Cohen's d), Correlation or Fisher's Z -transformed Correlation, Regression Slopes (Slopes), Risk Ratios, or whether the paper analysed the raw data or summary statistics directly (i.e., means/proportions) (which we label as raw/mean/proportion). For all other effects we labelled as 'other'. To determine whether effect sizes were likely to vary based on some underlying continuous variable or were dependent on the context in which traits were measured, we assessed whether authors: 1) acknowledged this explicitly in the paper or 2) accounted for it in their models. If it was not clear, two of us (DWAN & PP) discussed the paper, and made a decision. For example, trait means, such as ectothermic metabolic rate (i.e., raw/mean/proportion-based effect size), will vary depending on the temperature being measured, the magnitude of differences between two treatments (e.g, contrast-based effect sizes - SMD or lnRR) might depend on the temperature or dosage used in treatments, and the strength of relationship between two variables (e.g., correlation between behaviour and metabolism) might depend on the temperature context under which metabolic rate or behaviour were measured. Given the high heterogeneity in effects used in comparative physiology, we expected this to be common to most effect sizes, however, it has not been formally quantified.
